---
title: "TSA Final Forecasting"
author: "Yinan Ding, Jinxi Liu, Zhengqi Jiao"
date: "2023-04-11"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r, package, message=FALSE, warning=FALSE}
#Load/install required package here
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
library(dplyr)
library(zoo)
```

##Checking working directory
```{r}
getwd()
```

##Import Data

```{r, import data}
hourlyprice.raw <- read.csv("./Data/Day-Ahead_Zonal_Price_NYC_2010-2022.csv",stringsAsFactors = FALSE) # hourly LBMP of NYC from January 2010 to December 2022
#hourlyload.raw <- read.csv("./Data/NYC Load_Forecast_2018-2022.csv",stringsAsFactors = FALSE) # hourly LBMP of NYC from January 2018 to December 2022
#NGprice.raw <- read.csv("./Data/Henry_Hub_Natural_Gas_Spot_Price.csv",stringsAsFactors = FALSE) # Daily NG Price from January 2018 to December 2022
#REgeneration.raw <- read.csv("./Data/Monthly Renewable Generation_2018-2022.csv",stringsAsFactors = FALSE) # RE generation from January 2008 to December 2022

head(hourlyprice.raw, 10)
#head(hourlyload.raw, 10)
```


##Data Wrangling - hourly to daily
```{r}
class(hourlyprice.raw$Date)
class(hourlyprice.raw$DAM.Zonal.LBMP)


hourlyprice.raw <- hourlyprice.raw %>%
  separate(date, into = c("date", "time"), sep = " ")

hourlyprice <- hourlyprice.raw %>%
  select(date,DAM.Zonal.LBMP) %>%
  rename(Price = DAM.Zonal.LBMP)

hourlyprice.raw$date <- as.Date(hourlyprice.raw$date,format = "%m/%d/%Y")
  
dailyprice <- hourlyprice %>%
  filter( !is.na(hourlyprice$Price)) %>% 
  group_by(date) %>% 
  summarise(daily_mean_price = mean(Price))
  # slice(1:(n() - 1))
            

dailyprice.ts <- msts(dailyprice$daily_mean_price,
                      seasonal.periods = c(7,365.25),
                      start = c(2010,1,1)) 
# hourlyload.raw <- hourlyload.raw %>%
#   separate(Eastern.Date.Hour  , into = c("date", "time"), sep = " ") %>%
#   separate(time, into = c("hour", "minute"), sep = ":")
# 
# hourlyload <- hourlyload.raw %>%
#   select(date,hour,DAM.Forecast.Load) %>%
#   rename(Demand = DAM.Forecast.Load)
#   
#   
# dailyload <- hourlyload %>%
#   filter(!is.na(hourlyload$Demand)) %>% 
#   group_by(date) %>% # here we left column with minutes out
#   summarise( daily_mean_load = mean(Demand))%>%
#   slice(1:(n() - 1))
# 
# class(dailyload$date)
# dailyload$date <- as.Date(dailyload$date,format = "%m/%d/%Y")
# 
# str(dailyload)

ggplot(dailyprice, aes(x=date, y=daily_mean_price)) +
  geom_line() +
  ggtitle("Average Daily Electricity Price in NYC from 2010 to 2022");
            
# ggplot(dailyload, aes(x=date,y=daily_mean_load)) +
#   geom_line() +
#   ggtitle("Average Daily Electricity Load in NYC from 2018 to 2022")+
#   ylab("megawatthours")

```

## NG fix missing dates
```{r}
# NGprice.raw$Day <-as.Date(NGprice.raw$Day,format = "%m/%d/%Y")
# # Create a sequence of dates from the minimum to maximum date in the dataset
# date_seq <- data.frame(Day = seq(min(NGprice.raw$Day), max(NGprice.raw$Day), by = "day"))
# 
# library(dplyr)
# # Merge the sequence of dates with the original dataset using a left join
# complete_data <- left_join(date_seq, NGprice.raw, by = "Day")
# 
# # Create a zoo object from the 'value' column of the complete dataset
# zoo_data <- zoo(complete_data$Natural.Gas.Spot.Price.Dollars.per.Million.Btu)
# 
# # Fill the missing values with the average of the previous day and the next day
# zoo_data_filled <- na.approx(zoo_data)
# 
# # Convert the filled zoo object back to a data frame
# NGprice.fixed <- data.frame(Date = complete_data$Day, value = coredata(zoo_data_filled))
# 
# ggplot(NGprice.fixed, aes(x=Date,y=value)) +
#   geom_line() +
#   ggtitle("Daily Natural Gas Spot Price from 2018 to 2022")+
#   ylab("Dollars per Million Btu")
```

##Data Wrangling - daily to monthly
```{r}
# NGprice.daily <- NGprice.fixed %>%
#   rename(Price = value)
# NGprice.daily$Date <-as.Date(NGprice.daily$Date,format = "%m/%d/%Y")
# 
# 
# # Group the data by month and calculate the average price for each month
# NGprice.monthly <- NGprice.daily %>%
#   mutate(month = format(Date, "%Y-%m")) %>%
#   group_by(month) %>%
#   summarise(avg_price = mean(Price))
# 
# 
# Monthlyprice <- dailyprice %>%
#   mutate(month = format(date, "%Y-%m")) %>%
#   group_by(month) %>%
#   summarise(monthly_price = mean(daily_mean_price))
# 
# Monthlyload <- dailyload %>%
#   mutate(month = format(date, "%Y-%m")) %>%
#   group_by(month) %>%
#   summarise(monthly_load = mean(daily_mean_load))
# 
# REgeneration.monthly <- REgeneration.raw %>% 
#   select(Total) %>%
#   rename(REgeneration_all = Total)
# 
# start_date <- as.Date("2018-01-01")
# end_date <- as.Date("2022-12-01")
# monthly_dates <- seq(start_date, end_date, by = "month")
# 
# REgeneration.monthly <- REgeneration.monthly %>% 
#   mutate(date = monthly_dates)
# 
# ggplot(REgeneration.monthly, aes(x=date,y=REgeneration_all)) +
#   geom_line() +
#   ggtitle("Monthly REgeneration from 2018 to 2023")+
#   ylab("Dollars per Million Btu")

```

##Correlation Test
```{r}
# correlation.df <- data.frame(Monthlyprice$monthly_price,REgeneration.monthly$REgeneration_all, NGprice.monthly$avg_price,Monthlyload$monthly_load)
# 
# correlation.df <- correlation.df %>%
#   rename(Eprice=Monthlyprice.monthly_price,NGprice=NGprice.monthly.avg_price,REgeneration=REgeneration.monthly.REgeneration_all,Eload=Monthlyload.monthly_load)
# 
# correlation.model <- lm(data = correlation.df, Eprice~NGprice+REgeneration+Eload)
# summary(correlation.model)
# cor_matrix <- cor(correlation.df)
# cor_matrix
# 
# cor_table <- data.frame(cor_matrix)
# cor_table %>%
#   kable("html", align = "c", caption = "Correlation Coefficient Matrix") %>%
#   kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
#   column_spec(1, border_right = TRUE)
```

## Seperating into train and test subsets
```{r}
dailyprice.train.ts <- subset(dailyprice.ts,
                                   end = length(dailyprice.ts)-365) #Jan 1st 2010 to Dec 30st 2021
dailyprice.test.ts <- subset(dailyprice.ts,
                                   start = length(dailyprice.ts)-365) # Jan 1st 2022 to Dec 30st 2022
autoplot (dailyprice.train.ts)
autoplot (dailyprice.test.ts)


# dailyload.ts <- msts(dailyload$daily_mean_load,
#                       seasonal.periods = c(7,365.25),
#                       start = c(2018,1,1)) 
# 
# NGprice.fixed.ts <- msts(NGprice.fixed$value,
#                       seasonal.periods = c(7,365.25),
#                       start = c(2018,1,1))
# #subsetting
# dailyload.train.ts <- subset(dailyload.ts,
#                                    end = length(dailyload.ts)-365) #Jan 1st 2018 to Dec 30st 2021
# NGprice.fixed.train.ts <- subset(dailyload.ts,
#                                    end = length(NGprice.fixed.ts)-365) #Jan 1st 2018 to Dec 31st 2021

```


## Decomposing time series objects
```{r}
dailyprice.train.ts %>% mstl() %>% autoplot()
```

## Exponential Smoothing State Space Model
```{r, ETS Model}
ETS_fit <-  stlf(dailyprice.train.ts, h=365)

ETS_plot<-autoplot(dailyprice.test.ts, color = "dark grey") +
  autolayer(ETS_fit$mean, series="STL + ETS",PI=FALSE, color = "blue") +
  ylab("Electricity Price")
ETS_plot
```


## ARIMA with dynamic harmonic fourier components
```{r, ARIMA with dynamic harmonic fourier components}
ARIMA <- auto.arima(dailyprice.train.ts,
                  seasonal=TRUE, 
                  lambda=0,
                  xreg=fourier(dailyprice.train.ts,K=c(2,12)))
# ARIMA_load <- auto.arima(dailyprice.train.ts,xreg=fourier(dailyload.train.ts,K=c(2,12)))
# ARIMA_ng<- auto.arima(dailyprice.train.ts,xreg=fourier(NGprice.fixed.train.ts,K=c(2,12)))
# 
# dailyload.train.ts_fc <- forecast(dailyload.train.ts,NGprice.fixed.train.ts, h = 365,level=0.9)
# NGprice.fixed.train.ts_fc <- forecast(dailyload.train.ts, h = 365)

ARIMA_fc <- forecast(object = ARIMA,h=365)
# ARIMA_load_fc <- forecast(object = ARIMA_origin_fc,xreg=fourier(dailyload.train.ts_fc$mean,K=c(2,12),h=365),h=365)
# ARIMA_ng_fc <- forecast(object = ARIMA_origin_fc,xreg= fourier(NGprice.fixed.train.ts_fc$mean,K=c(2,12),h=365),h=365)

#Plot forecast with observed data
ARIMA_plot <- autoplot(dailyprice.test.ts, color = "dark grey") +
  autolayer(ARIMA_fc$mean, series="ARIMA_xgre",PI=FALSE,color = "blue") +
  ylab("Electricity Price Forecast")
ARIMA_plot
# 
# ARIMA_load_plot <- autoplot(dailyprice.test.ts, color = "dark grey") +
#   autolayer(ARIMA_load_fc$mean, series="ARIMA_xgre",PI=FALSE,color = "blue") +
#   ylab("Electricity Price Forecast with EXO(load)")
# ARIMA_load_plot 
# 
# ARIMA_ng_plot <-autoplot(dailyprice.test.ts, color = "dark grey") +
#   autolayer(ARIMA_ng_fc$mean, series="ARIMA_xgre",PI=FALSE,color = "blue") +
#   ylab("Electricity Price Forecast with EXO(load)")
# ARIMA_ng_plot
```

## TBATS
```{r, TBATS}
TBATS_fit <- tbats(dailyprice.train.ts)

#Forecast with TBATS
TBATS_fc <- forecast(TBATS_fit, h=365)

#Plot forecast with observed data
autoplot(dailyprice.test.ts, color = "dark grey") +
  autolayer(TBATS_forecast$mean, series="TBATS",PI=FALSE, color = "blue")+
  ylab("Electricity Price")
```


## Neural Network
```{r, Neural Network}
NN_origin <- nnetar(dailyprice.train.ts,p=1,P=0,xreg=fourier(dailyprice.train.ts, K=c(2,12)))
#NN_load <- nnetar(dailyprice.train.ts,p=1,P=0,xreg=fourier(dailyload.train.ts,K=c(2,12)))
#NN_ng <- nnetar(dailyprice.train.ts,p=1,P=0,xreg=fourier(NGprice.fixed.train.ts,K=c(2,12)))


#Forecast with NNet
NN_origin_fc<- forecast(NN_origin,h=365,xreg=fourier(dailyprice.train.ts,K=c(2,12),h=365))
#NN_load_fc <- forecast(NN_load, h=365,xreg=fourier(dailyload.train.ts_fc$mean,K=c(2,12)))
#NN_ng_fc <-forecast(NN_ng, h=365,xreg=fourier(NGprice.fixed.train.ts,K=c(2,12)))

#Plot forecast with observed data
NN_origin_plot <- autoplot(dailyprice.test.ts, color = "dark grey") +
  autolayer(NN_origin_fc$mean, series="Neural Network",PI=FALSE, color = "blue")+
  ylab("Electricity Price") 

# NN_load_plot <- autoplot(dailyprice.test.ts, color = "dark grey") +
#   autolayer(NN_load_fc$mean, series="Neural Network",PI=FALSE, color = "blue")+
#   ylab("Electricity Price") 
# NN_load_plot
# 
# NN_ng_plot <- autoplot(dailyprice.test.ts, color = "dark grey") +
#   autolayer(NN_ng_fc$mean, series="Neural Network",PI=FALSE, color = "blue")+
#   ylab("Electricity Price") 
# NN_ng_plot

```

## Checking accuracy of models

```{r}
#Model 1: STL + ETS
ETS_score <- accuracy(ETS_fit$mean, dailyprice.test.ts)  
#Model 2: ARIMA 
ARIMA_score <- accuracy(ARIMA_fc$mean, dailyprice.test.ts)
# #Model 2.1: ARIMA + Fourier(Load)
# ARIMA_load_score <- accuracy(ARIMA_load_fc$mean, dailyprice.test.ts)
# #Model 2.2: ARIMA + Fourier(NG)
# ARIMA_ng_score <- accuracy(ARIMA_ng_fc$mean, dailyprice.test.ts)

# Model 3:  TBATS 
TBATS_score <- accuracy(TBATS_fc$mean, dailyprice.test.ts)
# Model 4:  Neural Network 
NN_score <- accuracy(NN_origin_fc$mean, dailyprice.test.ts)
# # Model 4.1:  Neural Network 
# NN_load_score <- accuracy(NN_load_fc$mean, dailyprice.test.ts)
# # Model 4.2:  Neural Network 
# NN_ng_score <- accuracy(NN_ng_fc$mean, dailyprice.test.ts)

```


## Compare performance metrics
```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_score, ARIMA_score, TBATS_score, NN_score)
  )
row.names(scores) <- c("ETS", "ARIMA","TBATS", "NN")

# scores <- as.data.frame(
#   rbind(ETS_score, ARIMA_score,ARIMA_load_score,ARIMA_ng_score, TBATS_score, NN_score,NN_load_score, NN_load_score)
#   )
# row.names(scores) <- c("ETS", "ARIMA","ARIMA + exo(load)" ,"ARIMA + exo(ng)","TBATS", "NN", "NN+ exo(load)", "NN+ exo(ng)")
#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       

```


```{r}
kbl(scores, 
      caption = "Forecast Accuracy for Electricity Price",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(scores[,"RMSE"]))
```

```{r}
autoplot(dailyprice.test.ts) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_origin_fc, PI=FALSE, series="ARIMA") +
  autolayer(TBATS_fc,PI=FALSE, series="TBATS") +
  autolayer(NN_origin_fc,PI=FALSE, series="NN") +
  xlab("Day") + ylab("Electricity Price") +
  guides(colour=guide_legend(title="Forecast"))
```

## Predicting the future

```{r}
ETS_fit_2023 <-  stlf(dailyprice.ts, h=365)

autoplot(dailyprice.test.ts, color = "dark grey") +
  autolayer(ETS_fit_2023$mean, series="STL + ETS",PI=FALSE, color = "blue") +
  ylab("Electricity Price")
```

