---
title: "TSA Final Forecasting"
author: "Yinan Ding, Jinxi Liu, Zhengqi Jiao"
date: "2023-04-11"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r, package, message=FALSE, warning=FALSE}
#Load/install required package here
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(kableExtra)
library(dplyr)
```

##Checking working directory
```{r}
getwd()
```

##Import Data

```{r, import data}
library(readxl)
hourlyprice.raw <- read.csv("./Data/Day-Ahead_Zonal_NYC_2018-2022.csv",stringsAsFactors = FALSE) # hourly LBMP of NYC from January 2018 to December 2022
hourlyload.raw <- read.csv("./Data/NYC Load_Forecast_2018-2022.csv",stringsAsFactors = FALSE) # hourly LBMP of NYC from January 2018 to December 2022
NGprice.raw <- read.csv("./Data/Henry_Hub_Natural_Gas_Spot_Price.csv",stringsAsFactors = FALSE) # Daily NG Price from January 2018 to December 2022
REgeneration.raw <- read.csv("./Data/Monthly Renewable Generation_2018-2022.csv",stringsAsFactors = FALSE) # RE generation from January 2008 to December 2022

head(hourlyprice.raw, 10)
head(hourlyload.raw, 10)
```

##Data Wrangling - hourly to daily
```{r}
class(hourlyprice.raw$Date)
class(hourlyprice.raw$DAM.Zonal.LBMP)

hourlyprice.raw <- hourlyprice.raw %>%
  separate(Date, into = c("date", "time"), sep = " ") %>%
  separate(time, into = c("hour", "minute"), sep = ":")

hourlyprice <- hourlyprice.raw %>%
  select(date,hour,DAM.Zonal.LBMP) %>%
  rename(Price = DAM.Zonal.LBMP)

  
dailyprice <- hourlyprice %>%
  filter( !is.na(hourlyprice$Price)) %>% 
  group_by(date) %>% # here we left column with minutes out
  summarise( daily_mean_price = mean(Price))

class(dailyprice$date)
dailyprice$date <- as.Date(dailyprice$date,format = "%m/%d/%Y")
            
ggplot(dailyprice, aes(x=date,y=daily_mean_price)) +
  geom_line() +
  ylab("Average Daily Electricity Price in NYC from 2018 to 2022")

dailyprice.ts <- msts(dailyprice$daily_mean_price,
                      seasonal.periods = c(7,365.25),
                      start = c(2018,1,1)) 

autoplot(dailyprice.ts)

hourlyload.raw <- hourlyload.raw %>%
  separate(Eastern.Date.Hour  , into = c("date", "time"), sep = " ") %>%
  separate(time, into = c("hour", "minute"), sep = ":")

hourlyload <- hourlyload.raw %>%
  select(date,hour,DAM.Forecast.Load) %>%
  rename(Demand = DAM.Forecast.Load)
  
  
dailyload <- hourlyload %>%
  filter(!is.na(hourlyload$Demand)) %>% 
  group_by(date) %>% # here we left column with minutes out
  summarise( daily_mean_load = mean(Demand))

class(dailyload$date)
dailyload$date <- as.Date(dailyload$date,format = "%m/%d/%Y")

str(dailyload)
            
ggplot(dailyload, aes(x=date,y=daily_mean_load)) +
  geom_line() +
  ylab("Average Daily Electricity Price in NYC from 2018 to 2022")

```

## NG fix missing dates
```{r}
NGprice.raw$Day <-as.Date(NGprice.raw$Day,format = "%m/%d/%Y")
# Create a sequence of dates from the minimum to maximum date in the dataset
date_seq <- data.frame(Day = seq(min(NGprice.raw$Day), max(NGprice.raw$Day), by = "day"))

library(dplyr)
# Merge the sequence of dates with the original dataset using a left join
complete_data <- left_join(date_seq, NGprice.raw, by = "Day")

# Create a zoo object from the 'value' column of the complete dataset
zoo_data <- zoo(complete_data$Natural.Gas.Spot.Price.Dollars.per.Million.Btu)

# Fill the missing values with the average of the previous day and the next day
zoo_data_filled <- na.approx(zoo_data)

# Convert the filled zoo object back to a data frame
NGprice.fixed <- data.frame(Date = complete_data$Day, value = coredata(zoo_data_filled))
```

##Data Wrangling - daily to monthly
```{r}
NGprice.daily <- NGprice.fixed %>%
  rename(Price = value)
NGprice.daily$Date <-as.Date(NGprice.daily$Date,format = "%m/%d/%Y")


# Group the data by month and calculate the average price for each month
NGprice.monthly <- NGprice.daily %>%
  mutate(month = format(Date, "%Y-%m")) %>%
  group_by(month) %>%
  summarise(avg_price = mean(Price))
write.csv(NGprice.monthly, "NGprice.monthly.csv", row.names = FALSE)

Monthlyprice <- dailyprice %>%
  mutate(month = format(date, "%Y-%m")) %>%
  group_by(month) %>%
  summarise(monthly_price = mean(daily_mean_price))

Monthlyload <- dailyload %>%
  mutate(month = format(date, "%Y-%m")) %>%
  group_by(month) %>%
  summarise(monthly_load = mean(daily_mean_load))

REgeneration.monthly <- REgeneration.raw %>% 
  select(Month, Total) %>%
  rename(Date = Month, REgeneration_all = Total)

```

##Correlation Test
```{r}
correlation.df <- data.frame(Monthlyprice$monthly_price,REgeneration.monthly$REgeneration_all, NGprice.monthly$avg_price,Monthlyload$monthly_load)

correlation.df <- correlation.df %>%
  rename(Eprice=Monthlyprice.monthly_price,NGprice=NGprice.monthly.avg_price,REgeneration=REgeneration.monthly.REgeneration_all,Eload=Monthlyload.monthly_load)

correlation.model <- lm(data = correlation.df, Eprice~NGprice+REgeneration+Eload)
summary(correlation.model)
cor_matrix <- cor(correlation.df)
cor_matrix

cor_table <- data.frame(cor_matrix)
cor_table %>%
  kable("html", align = "c", caption = "Correlation Coefficient Matrix") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  column_spec(1, border_right = TRUE)
```

## Seperating into train and test subsets
```{r}
dailyprice.train.ts <- subset(dailyprice.ts,
                                   end = length(dailyprice.ts)-365) #Jan 1st 2018 to Dec 31st 2021
dailyprice.test.ts <- subset(dailyprice.ts,
                                   start = length(dailyprice.ts)-365) # Jan 1st 2022 to Dec 31st 2022
autoplot (dailyprice.train.ts)
autoplot (dailyprice.test.ts)


dailyload.ts <- msts(dailyload$daily_mean_load,
                      seasonal.periods = c(7,365.25),
                      start = c(2018,1,1)) 
#subsetting
dailyload.train.ts <- subset(dailyload.ts,
                                   end = length(dailyload.ts)-365) #Jan 1st 2018 to Dec 31st 202
```
```

## Decomposing time series objects
```{r}
dailyprice.ts %>% mstl() %>% autoplot()

```

```

## Exponential Smoothing State Space Model
```{r, ETS Model}
ETS_fit <-  stlf(dailyprice.train.ts, h=365)

autoplot(dailyprice.ts, color = "dark grey") +
  autolayer(ETS_fit$mean, series="STL + ETS",PI=FALSE, color = "blue") +
  ylab("Electricity Price")
```




## ARIMA with dynamic harmonic fourier components
```{r, ARIMA with dynamic harmonic fourier components}
ARIMA_fit <- auto.arima(dailyprice.train.ts,xreg=fourier(dailyload.train.ts,K=c(2,18))) 

#Forecast with ARIMA
dailyload.train.ts_for <- forecast(dailyload.train.ts, h = 365)
ARIMA_forecast <- forecast(object = ARIMA_fit,xreg=fourier(dailyload.train.ts_for$mean,K=c(2,18)),h=365)

#Plot forecast with observed data
autoplot(dailyprice.ts, color = "dark grey") +
  autolayer(ARIMA_forecast$mean, series="ARIMA_xgre",PI=FALSE, color = "blue") +
  ylab("Electricity Pricet")
```

## TBATS
```{r, TBATS}
TBATS_fit <- tbats(dailyprice.train.ts)

#Forecast with TBATS
TBATS_forecast <- forecast(TBATS_fit, h=365)

#Plot forecast with observed data
autoplot(dailyprice.ts, color = "dark grey") +
  autolayer(TBATS_forecast$mean, series="TBATS",PI=FALSE, color = "blue")+
  ylab("Electricity Price")
```

## Neural Network
```{r, Neural Network}
NN_fit <- nnetar(dailyprice.train.ts,p=1,P=0,xreg=fourier(dailyload.train.ts,K=c(2,18)))

#Forecast with NNet
NN_forecast <- forecast(NN_fit, h=365,xreg=fourier(dailyload.train.ts_for$mean,K=c(2,18)))

#Plot forecast with observed data
autoplot(dailyprice.ts, color = "dark grey") +
  autolayer(NN_forecast$mean, series="Neural Network",PI=FALSE, color = "blue")+
  ylab("Electricity Price") 
```

## Checking accuracy of models

```{r}
#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean, dailyprice.test.ts)  

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_forecast$mean, dailyprice.test.ts)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_forecast$mean, dailyprice.test.ts)

# Model 3:  Neural Network 
NN_scores <- accuracy(NN_forecast$mean, dailyprice.test.ts)

```

## Compare performance metrics
```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores)
  )
row.names(scores) <- c("ETS", "ARIMA", "TBATS", "NNet")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       

```

## Predicting the future

```{r}
ETS_fit_2023 <-  stlf(dailyprice.ts, h=365)

autoplot(dailyprice.ts, color = "dark grey") +
  autolayer(ETS_fit_2023$mean, series="STL + ETS",PI=FALSE, color = "blue") +
  ylab("Electricity Price")
```